\documentclass[12pt,a4paper]{article}
\oddsidemargin 0pt
\textwidth 160mm

\usepackage{fontspec}
\setmonofont[Path=/home/plain/.fonts/]{apl385.ttf}

\title{Executing APL code from inside \TeX}
\author{Dirk Laurie}
\date{14 March 2015}

\newcommand\showapl[1]{\directlua{tex.print(apl.texeval"#1")}}
\newcommand\doapl[1]{\directlua{tex.print(apl.texeval("#1",1))}}
\directlua{apl=require"gnuapl"}

\begin{document}
\maketitle

The \verb"gnuapl" Lua package allows you to evaluate APL code 
directly in \TeX\ and display it together with its output as it would 
appear in a GNU APL interactive session. No externally generated files, 
no cutting and pasting. All you need to do is execute
\begin{verbatim}
   lualatex myfile.tex
\end{verbatim}
for a suitable input file containing commands like 
\verb"\showapl{1⌽3 4⍴⍳12}". That command inserts the following output
into \verb"myfile.pdf".

\showapl{1⌽3 4⍴⍳12}

The macro \verb"\showapl" is defined in the preamble of \verb"myfile.tex" 
(see below), which also contains a command to load the package. In this 
document we describe only the \verb"texeval" command of \verb"gnuapl". 
The rest of the package is aimed more at standalone Lua applications and 
is documented elsewhere. 

\subsection*{\texttt{texeval(str[,option]})}
   Pass the APL code in \texttt{str} to the GNU APL interpreter for 
evaluation and optionally include the code and its result in the \TeX\
document.

\begin{description}
\item{\texttt{str}} A string containing APL code, which will be evaluated
regardless of what option is chosen.
\item{\texttt{option}}
   \begin{description}
      \item{\texttt{option=1}} Give the code to \TeX, indented as in an APL session.
      \item{\texttt{option=2}} Give the result to \TeX, not indented.
      \item{\texttt{option=3}} Both of the above. This is the default.
   \end{description}
\end{description}

\pagebreak

\section*{Preamble of this document}

You can use the following preamble in your own \TeX\ source, after
of course customizing some of the fields. Note the \TeX\ macros for
streamlined use of \texttt{texeval}.

\begin{verbatim}
\documentclass[12pt,a4paper]{article}
\oddsidemargin 0pt
\textwidth 160mm

\usepackage{fontspec}
\setmonofont[Path=/home/plain/.fonts/]{apl385.ttf}

\title{Executing APL code from inside \TeX}
\author{Dirk Laurie}
\date{14 March 2015}

\newcommand\showapl[1]{\directlua{tex.print(apl.texeval"#1")}}
\newcommand\doapl[1]{\directlua{tex.print(apl.texeval("#1",1))}}
\directlua{apl=require"gnuapl"}
\end{verbatim}

The font \texttt{apl385.ttf} is freely available in many places in the 
Internet. Any other fixed-space font that offers pleasing APL glyphs 
should work as well. In my opinion, APL code does not look good in
a variable-space font.

\end{document}
